---
title: "Start-up"
author: Colin Gillespie ([\@csgillespie](https://twitter.com/csgillespie))
date: "Jumping Rivers ([\\@jumping_uk](https://twitter.com/jumping_uk))"
output: ioslides_presentation
css: css/left.css
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Slides: Your Start-up file}
-->


# Start up

- This chapter deals with start up files

- It's quite easy to make an error and stop R working 
  
  - it's also very easy to fix, just delete the file

## The .Rprofile file

When R starts up it will run

1.  the site-wide R start up configuration file

2.  `.Rprofile` in the current working directory (the directory where R is
    started on the command-line)

3. If that file doesn't exist, it will then check your home directory

## The .Rprofile file

Determine your home area:

```{r, }
path.expand("~")
``` 

We can again test whether this file exists using 

```{r, }
file.exists("~/.Rprofile")
``` 

# Ten Minute Task

##  Ten Minute Task

- Use 
```{r, message=FALSE}
    path.expand("~")
```
to determine your home directory and
```{r, eval=FALSE}
    file.exists("~/.Rprofile")
``` 
to see if you have an `.Rprofile`

- Open `.Rprofile` and add the line
  
```{r, }
    message("Today is ", Sys.Date())
``` 

- Restart R.

# Some Examples

## Example .Rprofile settings

  - `options()` is a list that contains a number of default options

  - In my `.Rprofile` file, I have the line

```{r, eval=FALSE}
options(prompt="R> ", # Exciting new R> prompt
        digits = 4, # No. of digits displayed
        show.signif.stars = FALSE, # I hate stars
        fix_width = 88, # Useful under Linux
        Ncpus = 6L, # Parallel package installation
        continue = " ", # Hide the `+`
        mc.cores = 6L) # Parallel (see later)
```

## Setting the CRAN mirror

```{r, eval=FALSE}
r = getOption("repos")             
r["CRAN"] = "https://cran.rstudio.com/"
options(repos = r)
rm(r)
``` 

## Useful functions
  
  - I also load a couple of useful functions.

  - Be **very** careful about loading functions by default; code is no
    longer portable.
  

```{r, eval=TRUE}
setnicepar = function(mar=c(3,3,2,1), mgp=c(2,0.4,0), 
    tck=-.01, cex.axis=0.9, las=1, mfrow=c(1,1), ...) {
     par(mar=mar, mgp=mgp, tck=tck,
       cex.axis=cex.axis, las=las, mfrow=mfrow, ...)
     }
```   

## Useful functions

-  It is perfectly acceptable just to put your functions in your `.Rprofile`.

-  Downside: clutters up your work space.

-  If you run `rm(list = ls())`, your functions will be deleted. 

## Useful functions

- One neat trick is to combine hidden objects and environments.

- When an object name starts with `.`, by default it doesn't appear in
  the output of the `ls()` function

```{r, }
.obj = 1
ls()
``` 

## Useful functions

This concept also works with environments. In our `.Rprofile` file
we create a *hidden* environment

```{r, }
.env = new.env()
``` 

Then add our functions to the environment

```{r, }
.env$ht = function(d, n=6) rbind(head(d, n), tail(d,n))
```

 At the end of the `.Rprofile` file, we use the
`attach()`

```{r, message=FALSE}
attach(.env)
``` 
 
