---
title: "Code Profiling"
author: Colin Gillespie ([\@csgillespie](https://twitter.com/csgillespie))
date: "Jumping Rivers ([\\@jumping_uk](https://twitter.com/jumping_uk))"
output: ioslides_presentation
css: css/left.css
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Slides: Code Profiling}
%\VignetteEncoding{utf8}
-->

# Code Profiling
## Code Profiling

  * Often you will have working code, but simply want it to run faster. 
  
    * In some cases it’s obvious where the bottle neck lies. 
    
    * Sometimes you will guess, relying on intuition. 
    
    * Sometime you're wrong
    
## Rprof()

  * `Rprof()` is a built-in tool for profiling the execution of R expressions. 
  
  * At regular time intervals, the profiler stops the R interpreter, records the current function call stack, and saves the information to a file. 
  
    * Stochastic
  
    * Each time we run a function R, the conditions have changed

## Rprof()

  * Unfortunately `Rprof()` is not user friendly
  


# profvis()
## profvis()
```{r eval=FALSE}
library("profvis")
profvis({
  data(movies, package = "ggplot2movies") # Load data
  movies = movies[movies$Comedy == 1,]
  plot(movies$year, movies$rating)
  model = loess(rating ~ year, data = movies) # loess regression line
  j = order(movies$year)
  lines(movies$year[j], model$fitted[j]) # Add line to the plot
})
```

## Exercise: Monopoly

In Monopoly moving around the board is complicated by the fact that rolling a double 
(a pair of 1's, 2's, ..., 6's) is special:

  * Roll two dice (total1): total_score = total1;
  
  * If you get a double, roll again (total2) and total_score = total1 + total2;
  
  * If you get a double, roll again (total3) and total_score = total1 + total2 + total3;
  
  * If roll three is a double, Go To Jail, otherwise move total_score.

## Code

```{r, eval=FALSE}
## current is a value between 1 & 40
df = data.frame(d1 = sample(seq(1, 6), 3, replace = TRUE),
                d2 = sample(seq(1, 6), 3, replace = TRUE))

df$Total = apply(df, 1, sum)
df$IsDouble = df$d1 == df$d2
  
if (df$IsDouble[1] & df$IsDouble[2] & df$IsDouble[3]) {
  current = 11#Go To Jail
} else if (df$IsDouble[1] & df$IsDouble[2]) {
  current = current + sum(df$Total[1:3])
} else if (df$IsDouble[1]) {
  current = current + sum(df$Total[1:2])
} else {
  current = current + df$Total[1]
}
```


## Exercise: Monopoly

 * However the code is (relatively) slow when we simulate the game multiple times
 
 * Using `profvis()` can you pinpoint where this function is slow? 
 
    * But, it runs too quickly for `profvis()`, so
    
      - wrap the function in a loop
      
      - reduce the sampling interval, `interval = 0.005`
      
 * How would you speed it up?

```{r eval=FALSE}
vignette("profiling", package = "efficientTutorial")
```
