---
title: "Exercises: Common issues"
author: "Colin Gillespie"
output: rmarkdown::html_vignette
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Exercises: Common issues}
-->

## Exercise 1

In this example, we are going to investigate loading a large data frame. First, we'll
generate a large matrix of random numbers and save it as a csv file:

```{r practical1, echo=-2}
N = 1e5
N = 1e4
m = as.data.frame(matrix(runif(N), ncol=1000))
write.csv(m, file="example.csv", row.names=FALSE)
```
We can read the file the back in again using `read.csv()`
```{r}
dd = read.csv("example.csv")
```
To get a baseline result, time the `read.csv()` function call above, e.g. 

```{r eval=TRUE, results="hide"}
system.time(read.csv("example.csv"))
```

We will now look ways of speeding up this step.

1. Explicitly define the classes of each column using `colClasses` in
`read.csv()`, for example, if we have 1000 columns that all have data
type numeric, then:
    ```{r cache=TRUE,results='hide', tidy=FALSE}
    read.csv(file="example.csv", colClasses = rep("numeric", 1000))
    ```
1.  Use the `saveRDS()` and `readRDS()` functions:
    ```{r cache=TRUE, results="hide"}
    saveRDS(m, file = "example.RData")
    readRDS(file = "example.RData")
    ```
1. Compare the speed of `read_csv()` from the **readr** package to `read.csv()`

1. How does `fread()` from the **data.table** package compare to the other solutions?

Which of the above give the biggest speed-ups? Are there any downsides to using these
techniques? Do your results depend on the number of columns or the number of rows?


## Exercise 2

1. In this question, we'll compare matrices and data frames. Suppose we have a matrix, `d_m`
```{r}
##For fast computers
#d_m = matrix(1:1000000, ncol=1000)
##Slower computers
d_m = matrix(1:10000, ncol=100)
dim(d_m)
```
and a data frame `d_df`:
```{r}
d_df = as.data.frame(d_m)
colnames(d_df) = paste0("c", 1:ncol(d_df))
```

  * Using the following code, calculate the relative differences between selecting the first column/row of a data frame and matrix. 
    ```{r results='hide', tidy=FALSE}
    microbenchmark::microbenchmark(times = 1000, unit = "ms", # milliseconds
          d_m[1,], d_df[1,], d_m[,1], d_df[,1])
    ```
  Can you explain the result? Try varying the number of replications.

  * When selecting columns in a data frame, there are a few different methods. For example, 
    ```{r results='hide'}
    d_df$c10
    d_df[,10]
    d_df[,"c10"]
    d_df[,colnames(d_df) == "c10"]
    ```
    Compare these four methods.
  * Using the `object.size()` function compare the object size of a matrix and data data frame.


