---
title: "Exercises: Parallel programming"
author: "Colin Gillespie"
output: rmarkdown::html_vignette
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Exercises: Parallel programming}
-->

> Do you want in-house training? Please get in [contact](https://www.jumpingrivers.com/contact)


In the **efficientTutorial** package there is a function, `snake_ladders()`, that simulates a 
game of [snakes and ladders](https://en.wikipedia.org/wiki/Snakes_and_Ladders) (for a single player)^[This example
is based on the real game after a particularly long and boring session with my 5 year old son]:
```{r}
library("efficientTutorial")
snakes_ladders()
```
The return value is the number of rolls required to finish the game. We can simulate multiple games via

```{r eval = FALSE}
n = 10
results = numeric(n)
for (i in 1:n) {
  results[i] = snakes_ladders()
}
```
1. Simulate `n = 10000` games and create a boxplot

1. Rewrite the loop using `sapply()`
```{r, eval=FALSE}
results = sapply(1:n, function(i) snakes_ladders())
```
Is `sapply()` faster than a for loop?

1. Now run in parallel using `parSapply()`:
  * Load the **parallel** package
  * Create a cluster object using `cl = makeCluster(no_of_cores)`. To start set
  `no_of_cores = 2`
  * Rewrite `sapply()` as `parSapply()`. Remember that `parSapply` has an additional argument `cl`
  * Stop the cluster using `stopCluster()`

Time the different versions.

```{r echo=FALSE, eval=FALSE}
library("parallel")
cl  = makeCluster(2) 
parSapply(cl, 1:4, snakes_ladders)
stopCluster(cl)
```
1. If you using Linux or a Mac, take a look at `mclappy()`; similar to `parSapply()`,
but without clusters.

