---
title: "Exercises: Profiling"
author: "Colin Gillespie"
output: rmarkdown::html_vignette
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Exercises: Profiling}
-->

> Do you need statistical or software support for a grant or a proposal? 
> [Jumping Rivers](https://www.jumpingrivers.com) can help out.

In Monopoly moving around the board is complicated by the fact that rolling a double 
(a pair of 1's, 2's, ..., 6's) is special:

  * Roll two dice (total1): total_score = total1;
  * If you get a double, roll again (total2) and total_score = total1 + total2;
  * If you get a double, roll again (total3) and total_score = total1 + total2 + total3;
  * If roll three is a double, Go To Jail, otherwise move total_score.
The following piece of code simulates rolling dice in a game of Monopoly. 
```{r, echo = -1}
current = 1
## current is a number between 1 and 40
df = data.frame(d1 = sample(seq(1, 6), 3, replace = TRUE),
                d2 = sample(seq(1, 6), 3, replace = TRUE))
  
df$Total = apply(df, 1, sum)
df$IsDouble = df$d1 == df$d2
  
if (df$IsDouble[1] & df$IsDouble[2] & df$IsDouble[3]) {
  current = 11#Go To Jail
} else if (df$IsDouble[1] & df$IsDouble[2]) {
  current = current + sum(df$Total[1:3])
} else if (df$IsDouble[1]) {
  current = current + sum(df$Total[1:2])
} else {
  current = current + df$Total[1]
}
```

Using `profvis()` can you pinpoint where this code is slow? How would you speed it up?

### Hints {-}

The function runs too quickly for `profvis()`. So we need to call it multiple times
and reduce the sampling interval
```{r eval=FALSE}
profvis::profvis({
  for(i in 1:10000) {
    current = sample(1:40, 1)
    df = data.frame(d1 = sample(seq(1, 6), 3, replace = TRUE),
                    d2 = sample(seq(1, 6), 3, replace = TRUE))
    
    df$Total = apply(df, 1, sum)
    df$IsDouble = df$d1 == df$d2
    
    if (df$IsDouble[1] & df$IsDouble[2] & df$IsDouble[3]) {
      current = 11#Go To Jail
    } else if (df$IsDouble[1] & df$IsDouble[2]) {
      current = current + sum(df$Total[1:3])
    } else if (df$IsDouble[1]) {
      current = current + sum(df$Total[1:2])
    } else {
      current = current + df$Total[1]
    }
  }
  }, interval = 0.05)
```
Parts to target:

  * Using `seq` instead of `:` looks suspect
  * A data frame has overheads, why not try a matrix
  * Instead of `apply` what about `colSums`
  * Use scaler `AND` - `&&` instead of `&`


