---
title: "Code Profiling"
author: Colin Gillespie
output: ioslides_presentation
css: css/left.css
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Slides: Code Profiling}
%\VignetteEncoding{utf8}
-->

# Code Profiling
## Code Profiling

  * Often you will have working code, but simply want it to run faster. 
    * In some cases it’s obvious where the bottle neck lies. 
    * Sometimes you will guess, relying on intuition. 
    * Sometime you're wrong
    
## Rprof()

  * `Rprof()` is a built-in tool for profiling the execution of R expressions. 
  * At regular time intervals, the profiler stops the R interpreter, records the current function call stack, and saves the information to a file. 
    * Stochastic
    * Each time we run a function R, the conditions have changed

## Rprof()

  * Unfortunately `Rprof()` is not user friendly
  


# profvis()
## profvis()
```{r eval=FALSE}
library("profvis")
profvis({
  data(movies, package = "ggplot2movies") # Load data
  movies = movies[movies$Comedy == 1,]
  plot(movies$year, movies$rating)
  model = loess(rating ~ year, data = movies) # loess regression line
  j = order(movies$year)
  lines(movies$year[j], model$fitted[j]) # Add line to the plot
})
```

## Vignette

```{r eval=FALSE}
vignette("profiling", package = "efficientR")
```
