---
title: "Parallel computing"
author: Colin Gillespie ([\@csgillespie](https://twitter.com/csgillespie))
date: "`r Sys.Date()`"
output: ioslides_presentation
css: css/left.css
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Slides: Parallel computing}
-->

# Parallel computing
## Parallel computing

  * Over the last few years, the speed of a single CPU core has remained static 
  
```{r echo=FALSE,  warning=FALSE, fig.width=6, fig.height=4, out.width="70%", fig.align="center"}
local(source("code/08-hardware_cpu_speed.R", local=TRUE))
```
  

## Parallel computing

  * The `parallel` package comes with base R
    ```{r}
    library("parallel")
    ```
  * To detect the number of cores on your machine, use
    ```{r}
    detectCores()
    ```
    
__Quick Exercise__: How many cores do you have?

## Can my application be parallelised?    

  * An easy rule of thumb to determine if your `for` loop can be parallelised is
to ask the question, can the loop be written in reverse and still give the same
answer? 
  * For example
    ```{}
    for i in N:2
      a[i] = a[i-1] + b[i]
    ```
would result in a different `a` vector, or depending on how `a` has been
defined, could crash your program. 

## Can my application be parallelised?    

  * However this loop
    ```{}
    for i in 1:N
      a[i] = monteCarloSimulation()
    ```
results in identical results regardless of loop order

  
## The apply family

The most common parallel applications to be parallelized are replacements of `lapply`, `sapply`, and  `apply`

```{r eval=FALSE}
parLapply(cl, x, FUN, ...)
parApply(cl = NULL, X, MARGIN, FUN, ...)
parSapply(cl = NULL, X, FUN, ..., 
          simplify = TRUE, USE.NAMES = TRUE) 
```

```{r echo=FALSE}
mc = function(i, ...) runif(1)
N = 1
```

## Example: Monte-Carlo simulation

  * Suppose `mc` is a Monte-Carlo simulation function. Then we would have something like
    ```{r}
    results = numeric(N)
    for(i in 1:N)
      results[i] = mc(i)
    ```

## Example: Monte-Carlo simulation

  * Suppose `mc` is a Monte-Carlo simulation function. Then we would have something like
    ```{r}
    results = numeric(N)
    for(i in 1:N)
      results[i] = mc(i)
    ```
  * We could re-write this `for` loop as
    ```{r}
    results = sapply(1:N, mc)
    ```
  * Using `sapply` isn't faster than a standard `for` loop
  
## Example: Monte-Carlo simulation

  * To run in parallel is relatively straightforward. 
  * We make a `cluster` objects and specify the numer of threads to generate
    ```{r}
    cl = makeCluster(2)
    ```
  * Then change `sapply` to `parSapply`
    ```{r}
    results = parSapply(cl, 1:N, mc)
    ```
  * and clean up
    ```{r}
    stopCluster(cl)
    ```

## Exercise

```{r eval=FALSE}
vignette("parallel", package="efficientTutorial")
```
