---
title: "The byte compiler"
author: Colin Gillespie ([\@csgillespie](https://twitter.com/csgillespie))
output: ioslides_presentation
css: css/left.css
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Slides: The byte compiler}
-->
```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE, 
  fig.align="center",
  fig.pos="t"
)
```
# Byte compiler
## Byte compiler

 * The compiler package has been part of R since version 2.13.0 
    - It translates R functions into another language that can be interpreted by a very fast interpreter
 * Since R 2.14.0, all of the standard functions and packages in R will be pre-compiled into byte-code
 * Since R 3.4.0 JIT compilation has been enabled
 
 
## Byte compiler: the mean() function

```{r}
getFunction("mean")
```

note the `bytecode` line

## Byte compiler

  * We can compile our own R functions and obtain byte code version that may run faster. 

## Example: Bad mean

```{r}
mean_r = function(x) {
  total = 0
  n = length(x)
  for(i in 1:n)
    total = total + x[i]/n
  total
}
```

## Compiled version

```{r}
library("compiler")
cmp_mean_r = cmpfun(mean_r)
cmp_mean_r  
```


## Since 3.4.0 this is done automatically!

You need to be careful with timings!

## Benchmarks (R < 3.4.0)

```{r, eval=FALSE}
# Generate some data
x = rnorm(1000)
microbenchmark::microbenchmark(times = 10, unit = "ms", # milliseconds
          mean_r(x), cmp_mean_r(x), mean(x))
#> Unit: milliseconds
#>           expr   min    lq  mean median    uq  max neval cld
#>      mean_r(x) 0.358 0.361 0.370  0.363 0.367 0.43    10   c
#>  cmp_mean_r(x) 0.050 0.051 0.052  0.051 0.051 0.07    10  b 
#>        mean(x) 0.005 0.005 0.008  0.007 0.008 0.03    10 a  
```

## Benchmarks

```{r echo=FALSE,  warning=FALSE, fig.width=6, fig.height=4, out.width="70%", fig.align="center"}
local(source("code/02-byte_f1.R", local=TRUE))
```
  
## Package authors

If you create a package, then you automatically compile the package on installation by adding

```{r eval=FALSE}
ByteCompile: true
```
to the DESCRIPTION file

## Packages on CRAN

Most R packages installed using `install.packages()` are not compiled 
```{r, eval=FALSE}
byte = tools::CRAN_package_db()$ByteCompile
sum(!is.na(byte))/length(byte)*100
# [1] 3.291
```
    

## You will be compiled!

  * We can force packages to be compiled by starting R with the environment variable `R_COMPILE_PKGS`
  * Add `R_COMPILE_PKGS=3` to `~/.Renviron`

```{r eval=FALSE}
## Windows users need Rtools
install.packages("ggplot2", 
                 type = "source", 
                 INSTALL_opts = "--byte-compile") 
```



